\mytitle

### 2020-04-22 paper.tex
Finished edit of Section 2

In subsequent work~\cite{turner2014new,Zhang2014}, 
the WFLO problem was solved using \emph{exact}  
optimization methodologies that guarantee that the returned solution is optimal with respect
to the specified objective or is provably within a given bound of optimal.
%(i.e., maximize total expected power). 
The exact methods were shown to yield higher energy values, yet they often suffered from 
high computational cost and long runtimes. In practice, it is often desireable to strike a balance between optimality and computational requirements: fast sub-optimal solutions are useful when solving the wind farm design problem numerous times (e.g., when considering different numbers of turbines and various allowed locations), while optimal solutions yield the best possible placements (admitted by the mathematical model) when longer run-times are available.

In what follows, we provide a literature review that covers both exact and approximate methods for WFLO. We subsequently relate our 
work to both approaches via our proposed quadratic optimization framework for WFLO.\footnote{We review 
only literature that considers the same wind farm model as the one presented at the beginning of this section. For a broader literature review see~\cite{Zhang2014}.}

\paragraph{Exact Optimization Methods} 
The WFLO problem can be solved  
to optimality, using
techniques from operations research (e.g., mixed-integer linear programming (MILP), quadratic programming (QP), and constraint programming (CP))~\cite{Zhang2014,turner2014new,donovan2005wind}.
These methods are \emph{exact}, in the sense that they guarantee that the returned solution is indeed
the best solution one can obtain for the given problem. Donovan was the first to solve WFLO using a
MILP approach based on the LS energy expression~\cite{donovan2005wind}. In~\cite{turner2014new}, the authors
approximated the SS energy expression using a quadratic and a linear approximation. The two resulting methods performed well
compared to existing approximate and earlier exact solutions.
\todo{Add the NLM approach - explain how it becomes a different problem \cite{ulku2019new}}

The most recent attempt to solve the wind farm variation that we presented 
in this section is given in~\cite{Zhang2014}. The authors compared
previously existing MILP and QP models to their approach for directly optimizing the SS objective. To this end, they used constraint programming (CP), which 
is an optimization technique that does not make assumptions on the functional form of the objective.
However, applying CP turned out to be computationally intractable for the larger standard WFLO benchmarks.

In our work, we build upon the approach in~\cite{donovan2005wind}, yet instead of using a linear version of the LS objective, we 
model WFLO using a quadratic representation. The difference between our approach and the quadratic model
proposed in~\cite{turner2014new} is that our model is not an approximation of the SS objective, 
but an exact representation of the LS objective.

As we mentioned before, the main advantage of exact techniques
is that they provide guarantees on solution quality. However, their disadvantage is
that they may take a very long time to execute before finding an optimal solution. 
In fact, Section~\ref{sec:QUBO4WFLO} shows that WFLO is computationally hard, a result
that we have not yet encountered in the WFLO literature. Therefore,
when one considers solving multiple wind farm design problems, a fast solution is of essence,
and hence one may turn to approximate WFLO.

\paragraph{Approximate Methods} In contrast to exact methods, approximation algorithms often provide   
quick and sufficiently accurate solutions. Several works solved 
WFLO using evolutionary (genetic) approaches~\cite{MOSETTI1994105,gonzalez2010optimization,grady2005placement}. 
These methods use two principles (cross-over and mutation)
to change existing solutions and improve them over time. Evolutionary methods do not guarantee termination 
(i.e., they may run forever without achieving a quality criteria), and when they do terminate (in practical cases) they do
not guarantee optimal solutions. Hence, one must often
introduce non-quality related stopping criteria, 
thus compromising on the quality of the obtained solution~\cite{davis1991handbook}.
Furthermore, there is no natural way of introducing constraints into genetic methods, which
leads to a potentially costly feasibility check for every solution found.

The first attempt to address these limitations 
was made using a local search approach for WFLO~\cite{ozturk2004heuristic}. 
The solution does not guarantee optimality, 
yet it circumvents the termination and feasibility checking issues of the evolutionary
methods. The local search approach attempts to apply either move, remove or add actions given an incumbent 
feasible solution. When a turbine is moved, removed or added, the new solution is evaluated. The procedure terminates 
when a stopping criteria is reached.
The main limitation of~\cite{ozturk2004heuristic} is that it cannot perform an action 
that leads to an infeasible state. Therefore, it quickly converges to a sub-optimal solution (local optimum) 
that can potentially be far
from the globally optimal solution. This problem was experimentally demonstrated in~\cite{rivas2009solving}.

To overcome this limitation, the authors of~\cite{rivas2009solving} used simulated annealing (SA), 
a neighborhood search method that allows visits to infeasible solutions. 
This creates a well-connected solution space that leads the algorithm to escape 
local optima. The SA method was shown to be superior to~\cite{ozturk2004heuristic} 
and to the genetic method proposed in~\cite{grady2005placement}. A drawback of the SA approach 
is its ad-hoc nature. The various components of the algorithm
 must be tailored and geared towards WFLO. This means that small adjustments of the WFLO setting
 (e.g., adding noise considerations), would require major changes to the SA solution. 
Additional methods and experimental comparisons between various approximate algorithms are available in~\cite{samorani2013wind}.

In our work, we also use an SA approach (similarly to~\cite{rivas2009solving}). However,
our methodology is generic since it is based on a model and
on the ability of specialized hardware to solve (non-ad-hoc) quadratic problems.

>>>>

In subsequent work~\cite{turner2014new,Zhang2014}, 
the WFLO problem was solved using \emph{exact}  
optimization methodologies that guarantee that the returned solution is optimal with respect
to the specified objective or is provably within a given bound of optimal.
%(i.e., maximize total expected power). 
The exact methods were shown to yield higher energy values, yet they often suffered from 
high computational cost and long run-times. In practice, it is often desirable to strike a balance between optimality and computational requirements: fast sub-optimal solutions are useful when solving the wind farm design problem numerous times (e.g., when considering different numbers of turbines and various allowed locations), while optimal solutions yield the best possible placements (admitted by the mathematical model) when longer run-times are available.

In what follows, we provide a literature review that covers both exact and approximate methods for WFLO. We subsequently relate our 
work to both approaches via our proposed quadratic optimization framework for WFLO.\footnote{We review 
only literature that considers the same wind farm model as the one presented at the beginning of this section. For a broader literature review see Zhang et al.\ \cite{Zhang2014}.}

\paragraph{Exact Optimization Methods} 
The WFLO problem can be solved  
to optimality, using
techniques from operations research (e.g., mixed-integer linear programming (MILP), quadratic programming (QP), and constraint programming (CP))~\cite{Zhang2014,turner2014new,donovan2005wind}.
These methods are \emph{exact}, in the sense that they guarantee that the returned solution is indeed
the best solution one can obtain for the given problem. Donovan was the first to solve WFLO using a
MILP approach based on the LS energy expression~\cite{donovan2005wind}. Turner et al.\ \cite{turner2014new} approximated the SS energy expression using a quadratic and a linear approximation. The two resulting methods performed well
compared to existing approximate and exact techniques.
\todo{Add the NLM approach - explain how it becomes a different problem \cite{ulku2019new}}

The most recent attempt to solve the wind farm variation that we presented 
in this section is given in Zhang et al.\ \cite{Zhang2014}. The authors compared
previously existing MILP and QP models to their approach for directly optimizing the SS objective. To this end, they used constraint programming (CP), which 
is an optimization technique that does not make assumptions on the functional form of the objective.
However, applying CP turned out to be computationally intractable for the larger standard WFLO benchmarks.

In our work, we build upon the approach in Donovan~\cite{donovan2005wind}, yet instead of using a linear version of the LS objective, we 
model WFLO using a quadratic representation. %The difference between our approach and the quadratic model
%proposed in
Unlike Turner et al.\ \cite{turner2014new} our model is not an approximation of the SS objective, 
but an exact representation of the LS objective.

While the main advantage of exact techniques
is that they provide guarantees on solution quality, they may take a very long time to  find an optimal solution. 
In fact, Section~\ref{sec:QUBO4WFLO} shows that WFLO is computationally hard, a result
that we have not yet encountered in the WFLO literature. Therefore,
when one considers solving multiple wind farm design problems, a fast solution is of essence,
and hence one may turn to approximate WFLO.

\paragraph{Approximate Methods} In contrast to exact methods, approximation algorithms often provide   
quick and, hopefully, sufficiently accurate solutions. Several works solved 
WFLO using evolutionary (genetic) approaches~\cite{MOSETTI1994105,gonzalez2010optimization,grady2005placement}. 
These methods use two principles (cross-over and mutation)
to change existing solutions and improve them over time. Evolutionary methods do not guarantee termination 
(i.e., they may run forever without achieving a quality criteria), and even when they do terminate (in practical cases) they do
not guarantee optimal solutions. Hence, one must often
introduce non-quality related stopping criteria, 
thus compromising on the quality of the obtained solution~\cite{davis1991handbook}.
Furthermore, there is no natural way of introducing constraints into genetic methods, which
leads to a potentially costly feasibility check for every solution found.

The first attempt to address these limitations 
was made using a local search approach for WFLO~\cite{ozturk2004heuristic}. 
The solution does not guarantee optimality, 
yet it circumvents the termination and feasibility checking issues of the evolutionary
methods. The local search approach attempts to apply either move, remove or add actions given an incumbent 
feasible solution. When a turbine is moved, removed or added, the new solution is evaluated. The procedure terminates 
when a stopping criteria is reached.
The main limitation of this approach 
%~\cite{ozturk2004heuristic}
is that it cannot perform an action 
that leads to an infeasible state and, therefore, quickly converges to a sub-optimal solution (local optimum) 
that can potentially be much lower quality than the globally optimal solution~\cite{rivas2009solving}.

To overcome this limitation, Rivas et al.\ \cite{rivas2009solving} used simulated annealing (SA), 
a neighborhood search method that allows visits to infeasible solutions. 
This creates a well-connected solution space that leads the algorithm to escape 
local optima. The SA method was shown to be superior to Ozturk \& Norman~\cite{ozturk2004heuristic} 
and to the genetic method proposed in Grady et al.\ \cite{grady2005placement}. A drawback of the SA approach 
is its ad-hoc nature. The various components of the algorithm
 must be tailored to the problem at hand and, as a result, small adjustments of the WFLO setting
 (e.g., adding noise considerations), would require major changes to the SA implementation. 
Additional methods and experimental comparisons between various approximate algorithms are available in Samorani~\cite{samorani2013wind}.

In our work, we also use an SA approach (similarly to Rivas et al.\ \cite{rivas2009solving}). However,
our methodology is generic since it is based on a declarative problem model and
on the ability of specialized hardware to solve (non-ad-hoc) quadratic problems.
